How environmental drivers of spatial synchrony interact

### 2021-11-04 Paper/Paper.Rnw
various changes, catching up on forgetting to commit

%Alternative titles
%Interactions among Moran effects
%Interacting Moran effects
%Interactions between environmental drivers of spatial synchrony
%Environmental drivers of spatia synchony can strongly interact

@@

\begin{document}

@@

%Tentative paragraph outline written on my phone on 3/20/2021
%1) Importance of synchrony
%2) Classic problem of identifying causes of synchrony, but recently largely resolved due to new methods
%  -elaborate briefly on a couple methods, mention the importance of a timescale-resolved appproach
%3) in fact, it has recently been discovered that environmental causes of synchrony can interact, so that ...
%  -related to, but distinct from, interaction between environmental and dispersal causes of synchrony discussed, for instance by Kendal et al
%  -make sure you describe the two cases, cite the earlier papers, mention a few words about the two systems
%4) the simplest form of interacting moran effects can be demonstrated when a simple model, though elaboration of the true nature of this mechanism %will take the rest of this manuscript
%  -use Moran model
%  -but with two drivers
%  -moran theorem says correlation of pops equals correlation of sum of drivers, so work it out
%5) general idea of two noises lagged or time delayed relative to each other and their effects oppositely lagged - this may be where a pedagogical %fig comes in, or maybe it is just worsd here and then the pedagog fig comes in later, at end of theory section, in form of examples. Yes, I 
%think I want that second approach.
%6) brief paragraph on why we would expect interacting moran effects to be common, return to this point in the discussion
%7) final para on goals, manuscript structure, etc.
%  -G1) Elaborate a mathematical/statistical theory of interacting Moran effects which both allows both a detailed understanding 
%      of how the mechanism works and fecilitates statistical investigation of whether the effect is present in empirical systems.
%  -G2) Apply the theory and associated stats to kelp, the intention being to further elaborate the mechanism itself
%  -G3) Apply the theory and associated stats to plankton, the intention being to further elaborate the mechanism itself
%   Final sentence about how this paper conceptually and theoretically illuminates a new and probably common mechanism of 
%   synchrony, and provides statistics which facilitate investigation of the mechanism in empirical systems.

@@

%Tentative broad outline:
%DONE 1) Intro spectral quantities, the spectrum, the cospectrum, and the cross spectrum of two time series/processes. And the spectral
%matrix for putting them all in one structure.
%DONE 2) Define your measures of timescale/frequency-specific synchrony and crosss-variable synchrony, explain why these make sense as
%ways to measure those things.
%DONE 3) Present the model and the theory
%4) Present "toy" examples and fig 1 (need to think a lot about these)
%
%Notes:
%Compared to the deer paper, this is a much deeper presentation of the theory in the main text. I'm not sure that's avoidable.
%For one thing, the theory is explicitly applied to get the results, later, and that was not the case for the deer paper.

\noindent Our main theoretical work requires a conceptual understanding of the \emph{spectrum} of an environmental 
or population time series, as well as of the \emph{cospectrum} and \emph{cross spectrum} of two time series.
If $w_i(t)$ is a stochastic process or time series measured at location $i$ (e.g., the population density time series
of a species of interest in that location), the spectrum $S_{w_i w_i}(f)$ is a function
of frequency, $f$. For a periodic ocillation, $f$ can be defined as one over the timescale, or period, of the oscillation. 
The spectrum $S_{w_i w_i}(f)$ is larger for frequencies at which $w_i(t)$ oscillates more. So, for example, a 
population $w_i(t)$ that shows strong oscillatory dynamics of $5$-year period will have a large value of 
$S_{w_i w_i}(f)$ for $f=1/5 \text{ y}^{-1}$. The spectrum separates the overall variance of a time series into
contributions which occur on different frequencies, in the sense that an integral of $S_{w_i w_i}(f)$ across all 
frequencies equals $\var(w_i)$. In a similar way, the cospectrum of two time series, $w_i(t)$ and $w_j(t)$,
is a function of $f$ that takes large values at frequencies for which oscillations in $w_i(t)$ and 
$w_j(t)$ are both strong and strongly correlated, i.e., they have the same or similar phase. 
(Here, $j$ refers to another location where measurements were performed.) The cospectrum is 
the real part of the cross spectrum, $S_{w_i w_j}(f)$, which is a complex-valued function of frequency.
The cross-spectrum takes large-magnitude values at frequencies, $f$, for which the oscillatory components
of $w_i(t)$ and $w_j(t)$ are both strong; and the complex phase of the cross spectrum at $f$ then quantifies
the phase lag between those components. The notation $S_{ww}(f)$ refers to the \emph{spectral matrix}, 
the $ij^{\text{th}}$ entry of which is $S_{w_i w_j}(f)$. Spectral methods are standard tools in many fields, 
including ecology [e.g., \cite{Vasseur2007,Reuman2008,Defriez2017}], and many background references are available 
[e.g., \cite{Brillinger2001}].

Much prior work demonstrates the importance of a frequency-specific approach to
synchrony [e.g., \cite{Vasseur2007,Keitt2008,Defriez2016,Sheppard2016,Desharnais2018,Anderson2021}], 
and it will turn out (see below, and \emph{Discussion}) that a frequency-specific 
approach is essential to the development of our new theory. 
%***DAN: Good topic for a Discussion para, the fact that this could not really be done without taking frequency specificity into account, and why, though need to think a bit more before writing
We therefore here define, using the spectral methods outlined above, a frequency-specific measure of spatial synchrony, 
as well as a new concept of \emph{cross-variable spatial synchrony}. If time series data $w_i(t)$ for $t=1,\ldots,T$ 
were gathered at locations $i=1,\ldots,N$, our spatial synchrony measure is simply
$\rho_{ww}=\left( \sum_{i \neq j} S_{w_i w_j} \right)/(N^2-N)$, the average of the cross spectra for all pairs
of distinct locations. This is a real-valued function of frequency, the integral (across frequencies) of which
is the classic, non-frequency-specific measure $\left( \sum_{i \neq j} \cov(w_i,w_j) \right)/(N^2-N)$ (see SI section
\ref{SM-Sect:measuring_sync} for details, here). If two time series $\epsilon_i^{(1)}(t)$ and $\epsilon_i^{(2)}(t)$
($t=1,\ldots,T$ and $i,j=1,\ldots,N$) were measured at each sampling location (e.g., two environmental variables),
cross-variable spatial synchrony (or, simply, \emph{cross synchrony}) between the variables
is $\rho_{\epsilon^{(1)}\epsilon^{(2)}}=\left( \sum_{i \neq j} S_{\epsilon_i^{(1)}\epsilon_j^{(2)}} \right)/(N^2-N)$. 
This is conceptually interpretable as pertaining to spatial synchrony because it makes comparisons across 
distinct locations, i.e., $i \neq j$. 
It is interpretable as cross-variable synchrony because comparisons are between the two variables, i.e., cross spectra 
between time series components of $\epsilon^{(1)}$ and $\epsilon^{(2)}$ are used. The new index 
takes into account possible time lags. For instance, if both $\epsilon^{(1)}$ and $\epsilon^{(2)}$ show strong $4$-year-period 
oscillations, but peaks in the $\epsilon^{(2)}$ oscillations consistently lag peaks in the $\epsilon^{(1)}$ oscillations 
by a year, then $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$, which is complex valued, will have high magnitude at
frequency $f=1/4 \text{ y}^{-1}$, and will have  phase at that frequency equal to $\pi/2$, reflecting the one-year 
lag between the variables.
%***DAN: make sure I'm not making a sign error here when you create the additional examples for sup mat
See SI section \ref{SM-Sect:measuring_sync} for further examples of the new cross synchrony index.

Our model is 
\begin{align}
w_i(t) &= b_1 w_i(t-1)+ \cdots + b_n w_i(t-n) \label{main_model_1} \\ 
&+ p_0^{(1)} \epsilon_i^{(1)}(t)+ \cdots + p_{m_1}^{(1)} \epsilon_i^{(1)}(t-m_1) \label{main_model_2} \\
&+ p_0^{(2)} \epsilon_i^{(2)}(t)+ \cdots + p_{m_2}^{(2)} \epsilon_i^{(2)}(t-m_2) \label{main_model_3} \\
&+ \delta_i(t), \label{main_model_4}
\end{align}
where $w_i(t)$ is an index of the population in location $i=1,\ldots,N$ at time
$t$, and $\epsilon^{(1)}=(\epsilon_1^{(1)},\ldots,\epsilon_N^{(1)})$,
$\epsilon^{(2)}=(\epsilon_1^{(2)},\ldots,\epsilon_N^{(2)})$ and $\delta=(\delta_1,\ldots,\delta_N)$
are environmental processes. The processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$ are
taken to be measured environmental processes, whereas $\delta$ represents the 
aggregate influence of unmeasured processes. We assume that the combined process 
$(\epsilon_1^{(1)},\ldots,\epsilon_N^{(1)},\epsilon_1^{(2)},\ldots,\epsilon_N^{(2)},\delta_1,\ldots,\delta_N)$ is an ergodic second-order stationary stochastic process \citep{Brillinger2001}
with expected values of its component processes equal to $0$. We make additional mild
regularity assumptions to ensure the model is stable 
(SI section \ref{SM-sect:model_spec}). This model can straightforwardly be seen as 
a linearization of a very general dynamical model, influenced by ``weak noise'' [see, e.g.,
SI section S1.2 of \cite{Desharnais2018} and SI section S1 of \citep{Walter2017}].
Linearization and ``weak noise'' assumptions are tactics which have been made very commonly to make
theoretical progress in ecology (REFS), and it has been demonstrated that results based on a
weak noise assumption often also hold for quite strong noise (REF).
%***DAN: Add refs later
See \cite{Brillinger2001} for background on stochastic processes.

The outcome of our theory is an equation which expresses population spatial synchrony 
in terms of synchrony of the processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$, and cross
synchrony between these processes:
\begin{align}
\rho_{ww} &= \frac{1}{|f_{\mcB}|^2}\left[|f_{\mcP^{(1)}}|^2 \rho_{\epsilon^{(1)} \epsilon^{(1)}} 
+ |f_{\mcP^{(2)}}|^2 \rho_{\epsilon^{(2)} \epsilon^{(2)}} 
+ 2 \Rp \left(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}} \right)\right] \label{rhoww_T1to3} \\
&+ \text{other contributions}.\label{rhoww_OC}
\end{align}
Here, $f_{\mcB} = 1-b_1 \mu - b_2 \mu^2 - \cdots - b_n \mu^n$, 
$f_{\mcP^{(1)}} = p_0^{(1)} + p_1^{(1)} \mu + \cdots + p_{m_1}^{(1)} \mu^{m_1}$, and
$f_{\mcP^{(2)}} = p_0^{(2)} + p_1^{(2)} \mu + \cdots + p_{m_2}^{(2)} \mu^{m_2}$, where 
$\mu=\exp(-2 \pi \iota f)$, and where $\iota$, the Greek letter iota, represents the imaginary unit.
The ``other contributions'' in this equation correspond to synchronizing influences of 
$\delta$ and interactions between $\delta$ and the $\epsilon^{(i)}$. If data 
$w_i(t)$, $\epsilon^{(1)}_i(t)$ and $\epsilon^{(2)}_i(t)$ are available for locations 
$i=1,\ldots,N$ and times $t=1,\ldots,T$, then all the terms on the right side of (\ref{rhoww_T1to3})
can be estimated from these data (see \emph{Methods}), providing an assessment of the relative 
contributions of direct and interacting Moran effects to population synchrony.

%***DAN: There is still one thing left to do here, see above outline. This is a very important thing, because it is 
%supposed to use examples to demo how the interaction terms work, bringing in the lags in effects and the lags in
%the env vars themselves, and the interactions between these lags. This is where we take the next step with the 
%core of the concept, after the basic intro to it in the Intro.
%***DAN: Consider the possibility that you just want to use the same model as Plos Comp Biol, but analyze it using
%the new theory. On the other hand, don't want to give the impression that not much is new about this paper.

@@

@@

>>>>

%Alternative titles
%Interactions among Moran effects
%Interacting Moran effects
%Interactions between environmental drivers of spatial synchrony
%Environmental drivers of spatia synchony can strongly interact
%Environmental drivers of spatial synchrony interact

@@

\begin{document}
\SweaveOpts{concordance=TRUE}

@@

%Tentative paragraph outline written on my phone on 3/20/2021
%1) Importance of synchrony
%2) Classic problem of identifying causes of synchrony, but recently largely resolved due to new methods
%  -elaborate briefly on a couple methods, mention the importance of a timescale-resolved appproach
%3) in fact, it has recently been discovered that environmental causes of synchrony can interact, so that ...
%  -related to, but distinct from, interaction between environmental and dispersal causes of synchrony discussed, for instance by Kendal et al
%  -make sure you describe the two cases, cite the earlier papers, mention a few words about the two systems
%4) the simplest form of interacting moran effects can be demonstrated when a simple model, though elaboration of the true nature of this mechanism %will take the rest of this manuscript
%  -use Moran model
%  -but with two drivers
%  -moran theorem says correlation of pops equals correlation of sum of drivers, so work it out
%5) general idea of two noises lagged or time delayed relative to each other and their effects oppositely lagged - this may be where a pedagogical %fig comes in, or maybe it is just worsd here and then the pedagog fig comes in later, at end of theory section, in form of examples. Yes, I 
%think I want that second approach.
%6) brief paragraph on why we would expect interacting moran effects to be common, return to this point in the discussion
%7) final para on goals, manuscript structure, etc.
%  -G1) Elaborate a mathematical theory of interacting Moran effects which allows a detailed understanding 
%      of how the mechanism works 
%  -G2) Consider theoretical case studies, which emerge as important special cases of our general theory, and which illuminate the intuition behind how Moran effects may interact in real systems. 
%  -G3) Apply the theory to populations giant kelp of the California coast, the intention being to further elaborate the mechanism itself
%   Final sentence about how this paper conceptually and theoretically illuminates a new and probably common mechanism of 
%   synchrony.

@@

%Tentative broad outline:
%DONE 1) Intro spectral quantities, the spectrum, the cospectrum, and the cross spectrum of two time series/processes. And the spectral
%matrix for putting them all in one structure.
%DONE 2) Define your measures of timescale/frequency-specific synchrony and crosss-variable synchrony, explain why these make sense as
%ways to measure those things.
%DONE 3) Present the model and the theory
%4) Present "toy" examples and fig 1 (need to think a lot about these). DAN: perhaps these actually belong in Methods (describe the specific-case models we will consider) and results (apply the general theory to those specific cases)
%
%Notes:
%Compared to the deer paper, this is a much deeper presentation of the theory in the main text. I'm not sure that's avoidable.
%For one thing, the theory is explicitly applied to get the results, later, and that was not the case for the deer paper.

\noindent Our main theoretical work requires a conceptual understanding of the \emph{spectrum} of an environmental 
or population time series, as well as of the \emph{cospectrum} and \emph{cross spectrum} of two time series,
so we briefly introduce these concepts.
If $w_i(t)$ is a stochastic process or time series measured at location $i$ (e.g., the population density time series
of a species of interest in that location), the spectrum $S_{w_i w_i}(f)$ is a function
of frequency, $f$. For a periodic ocillation, $f$ can be defined as one over the timescale, or period, of the oscillation. 
The spectrum $S_{w_i w_i}(f)$ is larger for frequencies at which $w_i(t)$ oscillates more. So, for example, a 
population $w_i(t)$ that shows strong oscillatory dynamics of $5$-year period will have a large value of 
$S_{w_i w_i}(f)$ for $f=\frac{1}{5} \text{ y}^{-1}$. The spectrum separates the overall variance of a time series into
contributions which occur on different frequencies, in the sense that an integral of $S_{w_i w_i}(f)$ across all 
frequencies equals $\var(w_i)$. In a similar way, the cospectrum of two time series, $w_i(t)$ and $w_j(t)$,
is a function of $f$ that takes large values at frequencies for which oscillations in $w_i(t)$ and 
$w_j(t)$ are both strong and strongly correlated, i.e., they have the same or similar phase. 
(Here, $j$ refers to another location where measurements were performed.) The cospectrum is 
the real part of the cross spectrum, $S_{w_i w_j}(f)$, which is a complex-valued function of frequency.
The cross-spectrum takes large-magnitude values at frequencies, $f$, for which the oscillatory components
of $w_i(t)$ and $w_j(t)$ are both strong; and the complex phase of the cross spectrum at $f$ then quantifies
the phase lag between those components. The notation $S_{ww}(f)$ refers to the \emph{spectral matrix}, 
the $ij^{\text{th}}$ entry of which is $S_{w_i w_j}(f)$. Spectral methods are standard tools in many fields, 
including ecology [e.g., \cite{Vasseur2007,Reuman2008,Defriez2017}], and many background references are available 
[e.g., \cite{Brillinger2001}].

Much prior work demonstrates the importance of a frequency-specific approach to
synchrony [e.g., \cite{Vasseur2007,Keitt2008,Defriez2016,Sheppard2016,Desharnais2018,Anderson2021}], 
and it will turn out (see below, and \emph{Discussion}) that a frequency-specific 
approach is essential to the development of our new theory. 
%***DAN: Good topic for a Discussion para, the fact that this could not really be done without taking frequency specificity into account, and why, though need to think a bit more before writing
We therefore here define, using the spectral methods outlined above, a frequency-specific measure of spatial synchrony, 
as well as a new concept of \emph{cross-variable spatial synchrony}. If time series data $w_i(t)$ for $t=1,\ldots,T$ 
were gathered at locations $i=1,\ldots,N$, our spatial synchrony measure is simply
$\rho_{ww}=\left( \sum_{i \neq j} S_{w_i w_j} \right)/(N^2-N)$, the average of the cross spectra for all pairs
of distinct locations. This is a real-valued function of frequency, the integral (across frequencies) of which
is the classic, non-frequency-specific measure $\left( \sum_{i \neq j} \cov(w_i,w_j) \right)/(N^2-N)$ (see SI section
\ref{SM-Sect:measuring_sync} for details, here). If two time series $\epsilon_i^{(1)}(t)$ and $\epsilon_i^{(2)}(t)$
($t=1,\ldots,T$ and $i=1,\ldots,N$) were measured at each sampling location (e.g., two environmental variables),
cross-variable spatial synchrony (or, simply, \emph{cross synchrony}) between the variables
is $\rho_{\epsilon^{(1)}\epsilon^{(2)}}=\left( \sum_{i \neq j} S_{\epsilon_i^{(1)}\epsilon_j^{(2)}} \right)/(N^2-N)$. 
This is interpretable as pertaining to spatial synchrony because it makes comparisons across 
distinct locations, i.e., $i \neq j$. 
It is interpretable as cross-variable synchrony because comparisons are between the two variables, i.e., cross spectra 
between time series components of $\epsilon^{(1)}$ and $\epsilon^{(2)}$ are used. The new index 
takes into account possible time lags. For instance, if both $\epsilon^{(1)}$ and $\epsilon^{(2)}$ show strong,
spatially synchronous, $4$-year-period 
oscillations, but peaks in the $\epsilon^{(1)}$ oscillations consistently lag peaks in the $\epsilon^{(2)}$ oscillations 
by a year, then $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$, which is complex valued, will have high magnitude at
frequency $f=\frac{1}{4} \text{ y}^{-1}$, and will have  phase at that frequency equal to $\pi/2$, reflecting the one-year 
lag between the variables.
See SI section \ref{SM-sect:examples_cross_sync} for further examples of the new cross synchrony index.

Our model is 
\begin{align}
w_i(t) &= b_1 w_i(t-1)+ \cdots + b_n w_i(t-n) \label{main_model_1} \\ 
&+ p_0^{(1)} \epsilon_i^{(1)}(t)+ \cdots + p_{m_1}^{(1)} \epsilon_i^{(1)}(t-m_1) \label{main_model_2} \\
&+ p_0^{(2)} \epsilon_i^{(2)}(t)+ \cdots + p_{m_2}^{(2)} \epsilon_i^{(2)}(t-m_2) \label{main_model_3} \\
&+ \delta_i(t), \label{main_model_4}
\end{align}
where $w_i(t)$ is an index of the population in location $i=1,\ldots,N$ at time
$t$, and $\epsilon^{(1)}=(\epsilon_1^{(1)},\ldots,\epsilon_N^{(1)})$,
$\epsilon^{(2)}=(\epsilon_1^{(2)},\ldots,\epsilon_N^{(2)})$ and $\delta=(\delta_1,\ldots,\delta_N)$
are environmental processes. The processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$ are
taken to be measured environmental processes, whereas $\delta$ represents the 
aggregate influence of unmeasured processes. We assume that the combined process 
$(\epsilon_1^{(1)},\ldots,\epsilon_N^{(1)},\epsilon_1^{(2)},\ldots,\epsilon_N^{(2)},\delta_1,\ldots,\delta_N)$ is an ergodic second-order stationary stochastic process \citep{Brillinger2001}
with expected values of its component processes equal to $0$. We make additional mild
regularity assumptions to ensure the model is stable 
(SI section \ref{SM-sect:model_spec}). This model can straightforwardly be seen as 
a linearization of a very general dynamical model, influenced by ``weak noise'' [see, e.g.,
SI section S1.2 of \cite{Desharnais2018} and SI section S1 of \cite{Walter2017}].
Linearization and ``weak noise'' assumptions are tactics which have been used very commonly to make
theoretical progress in ecology (REFS), and it has been demonstrated that results based on a
weak noise assumption often also hold for noise which is quite strong (REF).
%***DAN: Add refs later
See \cite{Brillinger2001} for background on stochastic processes.

Fulfilling goal G1 of the Introduction (i.e., to elaborate theory of interacting Moran effects), 
the outcome of our theory is an equation which expresses population spatial synchrony 
in terms of synchrony of the environmenal processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$, and cross
synchrony between those processes:
\begin{align}
\rho_{ww} &= \frac{1}{|f_{\mcB}|^2}\left[|f_{\mcP^{(1)}}|^2 \rho_{\epsilon^{(1)} \epsilon^{(1)}} 
+ |f_{\mcP^{(2)}}|^2 \rho_{\epsilon^{(2)} \epsilon^{(2)}} 
+ 2 \Rp \left(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}} \right)\right] \label{rhoww_T1to3} \\
&+ \text{other contributions}.\label{rhoww_OC}
\end{align}
Here, $f_{\mcB} = 1-b_1 \mu - b_2 \mu^2 - \cdots - b_n \mu^n$, 
$f_{\mcP^{(1)}} = p_0^{(1)} + p_1^{(1)} \mu + \cdots + p_{m_1}^{(1)} \mu^{m_1}$, and
$f_{\mcP^{(2)}} = p_0^{(2)} + p_1^{(2)} \mu + \cdots + p_{m_2}^{(2)} \mu^{m_2}$, where 
$\mu=\exp(-2 \pi \iota f)$, and where $\iota$, the Greek letter iota, represents the imaginary unit.
The first term on the right of (\ref{rhoww_T1to3}) is the component of population spatial synchrony 
due to the direct Moran effects of $\epsilon^{(1)}$, the second term is 
the component of population synchrony due to the direct Moran effects of $\epsilon^{(2)}$, 
and the third term is the component due to interactions between the Moran effects of the 
two environmental drivers.
The ``other contributions'' in the above equations correspond to synchronizing influences of 
$\delta$ and of interactions between $\delta$ and the $\epsilon^{(i)}$. Such contributions would be
difficult to assess because $\delta$ was unmeasured. Comparing the three terms on the right
of (\ref{rhoww_T1to3}) gives the relative contributions of direct Moran effects and interactions 
between Moran effects. The detailed derivation of the theory is in SI section \ref{SM-sect:derivation}.

@@

\noindent We now describe how we pursue goal G2 of the Introduction, to develop 
theoretical case studies that illuminate the intuition of interacting Moran effects. 
We develop case studies A-C. For case study A,  the environmental variable 
$\epsilon^{(1)}$ is assumed to have a simple positive effect on populations, but 
lagged by one model time step, i.e.,
$p_0^{(1)}=0$, $p_1^{(1)}>0$, and $p_i^{(1)}=0$ for $i>1$. 
For case study B, which generalizes A, $\epsilon^{(1)}$ is again assumed to have a simple 
positive effect on populations,
but now lagged by $n$ time steps, so that $p_n^{(1)}>0$ while $p_i^{(1)}=0$ for $i \neq n$.
For both A and B, the effects of $\epsilon^{(2)}$
are assumed to be un-lagged, i.e., $p_0^{(2)} \neq 0$ and $p_i^{(2)}=0$ for $i>0$.
For case study C, the effects of $\epsilon^{(1)}$ are lagged by $n_1$ and the 
effects of $\epsilon^{(2)}$ are lagged by $n_2$, i.e.,
$p_{n_1}^{(1)}>0$ while $p_i^{(1)}=0$ for $i \neq n_1$, and 
$p_{n_2}^{(2)}>0$ while $p_i^{(2)}=0$ for $i \neq n_2$.
The noise process
$(\epsilon^{(1)},\epsilon^{(2)})$ is assumed to be a Gaussian white-noise process 
for both A and B, i.e., 
the random variables $(\epsilon^{(1)}(t),\epsilon^{(2)}(t))$ are independent and 
identically distributed (iid) for distinct times, $t$. Noise may be correlated across space,
and the components of $\epsilon^{(1)}(t)$ may be correlated with those of $\epsilon^{(2)}(t)$.
For C, the noise processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$ are
each assumed to exhibit spatially synchronous periodic oscillations of the same period
but potentially different phases. Such a scenario could be realized by seasonally 
periodic environmental fluctuations if the model time step were quarterly or monthly,
e.g., kelp wave action in central California peaks annually in the winter, whereas surface-water
nitrate concentrations also fluctuate with period one year, but peak in the spring. 
Details of the noise processes are specified in SI section \ref{SM-sect:examples_cross_sync}.
For all case studies,
the autoregressive order of population dynamics is assumed to be $1$, i.e, $b_1 \neq 0$ 
but $b_i=0$ for $i>1$.

@@

% What do we learn from case study A:
% 
% Warmup:
% 1) interactions between Moran effects (ref to specific term of the theory) can easily be of comparable importance
% to the main Moran effects
% 2) interaction effects can be positive or negative, and can differ, and even differ in sign, by timescale
% 
% Actually breaking it down:
% 1) At the highest frequencies (shortest timescales) of 0.5 cycles per sampling 
% interval (period 2 sampling intervals), a lag of 1 
% sampling interval in the population effects of \epsilon^{(1)} is half the period of oscialltion, so 
% causes the effects of \epsilon^{(1)}
% on those timescales to be exactly out of phase of the effects of \epsilon^{(2)}, and so we get 
% cancellation, i.e., negative interaction effects. The effects of \epsilon1 are multiplied by 1.1,
% those of \epsilon2 by 0.8, and the correlation between the two (which is made to have negative effects
% by the lag) is 0.3. So the negative interaction effects are 1.1*0.8*0.3*2, where the last factor of
% two is because there are actually two effects here, the one where \epsilon1 is operating, lagged, at i and 
% \epsilon2 is operating at j; and the effect where \epsilon1 is operating, lagged, at j and \epsilon2 is 
% operating at i. 
% 2) As timescales get to 1 year = 4 quarters, a lag of 1 quarter in the population effects of 
% \epsilon1 puts the population effects of \epsilon1 3/4 out of phase with those of \epsilon 2,
% and the correlation of two sin waves of the same frequency that are 3/4 out of phase is 0, so 
% interaction effects are 0 here.
% 3) As timescales get long, a lag of 1 quarter in the population effects of \epsilon1 is a
% negligible time compared to the timescale being considered, so the effects of 
% \epsilon1 and \epsilon2 on populations are closer and closer to be in phase and correlated
% as timescales get longer and longer. The effects of \epsilon1 are multiplied by 1.1, those of 
% \epsilon2 by 0.8, and the correlation is 0.3, so the interaction effect is 1.1*.8*.3*2 (the 
% reason for the 2 is explained above).
% 
% Need to put this into good writing
% 
% Need to carry out similar thoughts for case B and write them, too.

%Thoughts on 2021 11 01
% Topic sentence for first para on theory results:
% Results confirm earlier empirical work (refs) that interaction effects between Moran drivers 1) can be comparable in strength to direct Moran effects; 2) can be synergistic or destructive; and 3) can depend strongly on timescale, even to the extent that the sign of interaction effects can differ by timescale.
% 
% Combine all three theory figs into one multi panel fig, change the order of panels again 
% 
% Subsequent paras will then break things down to understand how this stuff works
% 
% Possible topic sentence for subsequent para: Theory reveals, as illustrated by cases A and B, how interactions between Moran effects depend on the lags with which Moran drivers influence populations. Or theory reveals how interactions between Moran effects depend on the lags with which Moran drivers influence populations, and case study begins to illustrate how this works. Or something like that. 
% 
%The also have a para for C
%
% Maybe get rid of the “confirm earlier work” phrase since it wasn’t necessarily explicitly listed in that way in earlier work

%Para basically confirming earlier work, as a starter
Theoretical results help demonstrate that interaction effects between Moran drivers: 1) can be 
comparable in strength to direct Moran effects; 2) can be synergistic or destructive; and 3) 
can depend strongly on timescale, even to the extent that the sign of interaction effects can 
differ by timescale.

%Current (2021 11 01) thoughts on figures:
%1) Maybe a pedagog fig, not sure what this would look like
%2) Combined theory fig, all 3 case studies
%3) "main fig" for CC1 and SB systems
%4-5) "explain direct effects" fig for CC1 and SB systems (2 figures, I think it'll have to be)
%6-7) "explain indirect effects" figs for CC1 and SB systems (again, 2)
%One too many figs here. What should be cut? Right now I don't know what the pedagog fig would be 
%so that's the obvious one to cut. And then it will all fit, though the figs are really big and 
%it seems like it is always good to have a pedagogical, warmup fig. If you do come up with a pedagog
%fig idea, could maybe combine some figs to make it work, or put the pedagog content as the first
%part of the theory fig? Don't worry about it for now.

%***DAN: add case study C to this.
\begin{figure}
\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseA.pdf}\\
\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseB.pdf}\\
\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseC_v01.pdf}\\
\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseC_v02.pdf}
\caption{Theoretical case studies A (panels a-d), B (panels e-h) and C (i-p, two alternative sets of 
parameters used for i-l and m-p, respectively). Left panels (a, e, i, m) show the terms on the right side
of (\ref{rhoww_T1to3}). On those panels, the green line is 
$\frac{|f_{\mcP^{(1)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(1)} \epsilon^{(1)}}$, quantifying 
population synchrony due to the direct Moran effects of $\epsilon^{(1)}$. The blue line
is $\frac{|f_{\mcP^{(2)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(2)} \epsilon^{(2)}}$, quantifying 
population synchrony due to the direct Moran effects of $\epsilon^{(2)}$.
The dashed green-blue line is 
$\frac{2 \Rp \left(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}} \right)}{|f_{\mcB}|^2}$, 
which is population synchrony due to interacting Moran effects.
%***DAN: More text here as we flesh out the presentation 
Model parameters were: $N=5$ and $b_1=0.4$ for all case studies; 
$p_1^{(1)}=1.1$ and $p_0^{(2)}=0.8$ for case study A; 
$p_3^{(1)}=1.1$ and $p_0^{(2)}=0.8$ for case study B;
$p_1^{(1)}=3$, $p_0^{(2)}=1.5$, $c_1=0$ and $c_2=-4/9$ for case study C;
$v_{11}=v{22}=0.6$ and $v_{12}=0.3$ for case studies A and B;
$w_{11}=w{22}=0.6$ and $w_{12}=0.3$ for case study C;
$l_1=0$ and $l_2=1$ for case study C, panels i-l; and 
$l_1=1$ and $l_2=0$ for case study C, panels m-p.
See Methods and SI sections \ref{SM-sect:examples_cross_sync} and 
\ref{SM-sect:details_case_studies} for the meaning of parameters and additional details.}\label{fig:TheoryCaseA}
\end{figure}
%#Case A
%N<-5
%ns<-c(.6,.6,.3)
%p1<-c(0,1.1)
%p2<-c(0.8)
%b<-c(.4)
% #Case B
% N<-5
% ns<-c(.6,.6,.3)
% p1<-c(0,0,0,1.1)
% p2<-c(0.8)
% b<-c(.4)
